---
title: "Homework 2"
author: "Wenzhao Wu"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
library(dbplyr)
```

## Problem 1

Read the Mr. Trashwheel dataset.

```{r}
trashwheel_df = 
  read_xlsx(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "Mr. Trash Wheel",
    range = cell_cols("A:N")) %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```


Read precipitation data!

```{r}
precip_2018 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
  ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2018) %>%
  relocate(year)


precip_2017 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2017 Precipitation",
    skip = 1
  ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2017) %>%
  relocate(year)
```

Now combine annual precipitation.

```{r}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df = 
  bind_rows(precip_2018, precip_2017)

left_join(precip_df, month_df, by = "month")
```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stroes it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include month precipitation data. 


## Problem 2

```{r}
nyc_transit_df = 
  read_csv(
    "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    col_types = cols(
      Route8 = col_character(),
      Route9 = col_character(),
      Route10 = col_character(),
      Route11 = col_character()
    )) %>%
  janitor::clean_names() %>%
  select(line:entry, vending, ada) %>%
  mutate(
    entry = recode(entry, "YES" = "TRUE", "NO" = "FALSE"),
    entry = as.logical(entry, levels = c("TRUE","FALSE")))

nyc_transit_df
```

This dataset contains data about line, stations' names, geographical locations and routes served. Also, it includes information of the entrance type, whether there exists an entry or vending, and whether the ADA is complied with. 
By far, I have cleaned up variable names to lower_snake_case, and selected only the columns of variables that I am interested in. 

There are `r nrow(nyc_transit_df)` rows and `r ncol(nyc_transit_df)` columns of the resulting dataset. These data are currently not tidy at all, because for example, some columns could have been combined together for tidiness. 

There are `r nrow(distinct(nyc_transit_df, line, station_name))` distinct stations shown in the dataset. 

In addition, there are `r filter(nyc_transit_df, ada == "TRUE") %>% distinct(line, station_name) %>% count()` distinct stations compliant with ADA.

The number of entrances without vending is `r nrow(filter(nyc_transit_df, vending == "NO"))`. Among these entrances, there are `r nrow(filter(nyc_transit_df, vending == "NO" & entry == "TRUE"))` allowing entry. Thus, the proportion of entrances without vending that allow entry would be `r 100*(nrow(filter(nyc_transit_df, vending == "NO" & entry == "TRUE"))) / (nrow(filter(nyc_transit_df, vending == "NO")))` %.


Reformatting data.

```{r}
nyc_tidy = 
  pivot_longer(
  nyc_transit_df,
  route1:route11,
  names_to = "route_names",
  values_to = "route_numbers",
  names_prefix = "route") %>%
  distinct(station_name, line, route_names, route_numbers, .keep_all = TRUE)
```

For distinct stations, there are `r nrow(filter(nyc_tidy,route_numbers == "A"))` serving the A train, of which there are `r nrow(filter(nyc_tidy, route_numbers == "A") %>% filter(ada == "TRUE"))` ADA compliant.


## Problem3

```{r pols}
pols_month = 
  read_csv(
    "./data/pols-month.csv") %>%
  janitor::clean_names()

pols_month =
  separate(pols_month, mon, into = c("year", "month_num", "day")) %>%
  mutate(
    year = as.integer(year),
    month_num = as.integer(month_num),
    day = as.integer(day)) %>%
  select(-prez_gop,-prez_dem,-day) %>%
  pivot_longer(
  gov_gop:rep_dem,
  names_to = "president",
  values_to = "pol")

month_df = 
  tibble(month_num = 1:12,
          month = month.abb)
joina_df = left_join(month_df,pols_month, by = "month_num")
pols_df = select(joina_df, -month_num) %>%
  relocate(year, month)
```

```{r snp}
snp = 
  read_csv(
    "./data/snp.csv") %>%
  janitor::clean_names()

snp =
  separate(snp, date, into = c("month_num", "day","year")) %>%
  mutate(
    year = as.integer(year),
    month_num = as.integer(month_num),
    day = as.integer(day)) %>%
  select(-day)

month_df = 
  tibble(month_num = 1:12,
          month = month.abb)

joinb_df = left_join(month_df,snp, by = "month_num")

snp_df = select(joinb_df,-month_num) %>%
  relocate(year, month)
```

```{r unemployment}
unemployment = 
  read_csv(
    "./data/unemployment.csv") %>%
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "unemp") %>%
  janitor::clean_names()
```
Merge 3 datasets.

```{r merged}
snp_pols = 
  left_join(snp_df,pols_df, by = c("year","month"))

all_df = 
  left_join(unemployment, snp_pols, by = c("year","month"))

all_df
```

The pol_df dataset contains information about presidents and their corresponding "pols" integers in each month from `r min(pull(pols_df,year))` to `r max(pull(pols_df, year))`. I used function **separate()** to break "mon" into three different variables. The columns of "prez_dem", "prez_gop" and "day" were removed by using **select()**. Then other "gop" and "dem" columns were combined by function **pivot_longer()**.

The snp_df includes data of `r names(snp_df)`, from `r min(pull(snp_df, year))` to `r max(pull(snp_df, year))`. The first two datasets were both replaced month numbers with month names by creating a separate dataframe which was then merged into the dataset using **left_join()**.

The third dataset of unemployment has `r nrow(unemployment)` rows and `r ncol(unemployment)` columns. The data ranges from `r min(pull(unemployment,year))` to `r max(pull(unemployment, year))`.

The merged dataset with combined information from the previous three datasets, which contains `r nrow(all_df)` rows and `r ncol(all_df)` columns. Variables includes `r names(all_df)`. The merging is based on the common columns "year" and "month", by the function **left_join**.



